Viscous Erosion in Porous Media

### 2018-04-04 docs/methods/erosion_methods.tex
NM

To further quantify our observations, Fig.~\ref{fig:mbodyplots} shows several measurements taken from the multibody simulations. In all cases, the black curves correspond to the 6-body simulation and gray curves to the 50-body simulation from Fig.~\ref{fig:50bodies}. First, Fig.~\ref{fig:mbodyplots}(a) shows the fraction of area occupied by the solid bodies as it varies in time. In the 6-body simulation, the initial configuration comprises about 40\% of the domain area, and this percentage steadily decreases as the bodies erode. The 50-body simulation shows a similar trend. Figure \ref{fig:mbodyplots}(b) shows how the flow rate, $\umax$, changes in time as a result of fixing the pressure drop across the cell. Notice that, in this figure, the vertical axis is logarithmic. Both simulations show the flow rate to increase rapidly, even during the early stages of erosion when the area fraction has not changed that significantly. In fact, due to the logarithmic axis, the nearly linear trend seen in the figures corresponds to exponential growth of $\umax$. The change in area-fraction alone is not enough to account for this rapid growth, implying that it is the way in which erosion reshapes the bodies that allows for high flow rates.

@@

Figure \ref{fig:mbodyplots}(b) shows the inlet/outlet flow-rate, $\umax$, over time. Since we have fixed the pressure drop across the cell in these simulations, $\umax$ steadily increases as the bodies diminish and provide less resistance. Here, the vertical axis is logarithmic, and so the plot suggests exponential increase of $\umax$ at early times. This rapid increase in flow is likely due to the erosion-induced channelization observed above. In Fig.~\ref{fig:mbodyplots}(c) we show the total drag, summed over all bodies, versus time. Though the bodies diminish due to erosion, the steady increase of $\umax$ causes the total drag to remain relatively constant. Only near the end of the simulation does the drag decrease rapidly to zero. To remove the effect of increasing flow rate, we show in Fig.~\ref{fig:mbodyplots}(d) the total drag scaled on $\umax$. This value is a measure for the aggregate {\em resistance} of all bodies. Due to linearity of the Stokes equations, this measure does not depend on the flow-rate $\umax$, only on the bodies and their configuration. As seen here, the resistance decreases rapidly as the bodies erode. Since the vertical axis is logarithmic, the plot suggests an exponential decrease in resistance at early times, which is likely connected to the channelization process.

In Fig.~\ref{fig:mbodyplots} we also plot the same quantities for the 50-body simulation that was shown in Fig.~\ref{fig:50bodies} of the introduction. Overall, these plots support the same trends seen in the 6-body simulation. The initial area-fraction of the 50-body simulation is only slightly higher, about 45\%, and this fraction decreases steadily as bodies erode. The flow rate $\umax$ again increases very rapidly with time, though with a significantly smaller initial value due to the higher resistance provided by 50 bodies. As before, the total drag remains relatively constant throughout most of the simulation and then abruptly decreases to zero in the final moments. Our measure of resistance, $F_D/(4 \pi \umax)$, again decreases very rapidly in the early moments when channels are initially formed. Though the area fractions of the two simulations are nearly the same, the 50-body configuration provides about 5 times more resistance.

>>>>

To further quantify these observations, Fig.~\ref{fig:mbodyplots} shows several measurements taken from the multibody simulations. In all cases, the black curves correspond to the above 6-body simulation and gray curves to the 50-body simulation from Fig.~\ref{fig:50bodies}. First, Fig.~\ref{fig:mbodyplots}(a) shows the fraction of area occupied by the solid bodies as it varies in time. In the 6-body simulation, the initial configuration comprises about 40\% of the domain area, and this percentage steadily decreases as the bodies erode. The 50-body simulation shows a similar trend.

Next, Figure \ref{fig:mbodyplots}(b) shows how the flow rate, $\umax$, changes in time as a result of fixing the pressure drop across the cell. Notice that, in this figure, the vertical axis is logarithmic. Both simulations show the flow rate to increase rapidly, even during the early stages of erosion when the area fraction has not changed that significantly. In fact, due to the logarithmic axis, the nearly linear trend seen in the figures corresponds to exponential growth of $\umax$. The change in area-fraction alone cannot account for this rapid growth, suggesting the reshaping process to be vital. In particular, the formations of channels observed in Fig.~\ref{fig:06bodies} is most likely the key feature that enables the higher flow rates.

Naturally, the flow rate allowed by the bodies is connected to the amount of drag that they exert. We therefore show in Fig.~\ref{fig:mbodyplots}(c) the total drag exerted by all bodies as it varies in time. In both cases, the drag remains nearly constant throughout much of the simulation and then drops to zero in the final moments. Here, the rapid increase of $\umax$ offsets the reduction in body size to create the nearly constant drag.

To remove this effect of increasing flow rate, we show in Fig.~\ref{fig:mbodyplots}(d) the drag by $\umax$. Since both the pressure and viscous drag depend linearly on $\umax$, this normalized quantity, $F_D / (4 \pi \umax)$, is {\em independent} of $\umax$ and therefore isolates how drag depends on the configuration of bodies. In this way, we can interpret $F_D / (4 \pi \umax)$ as the {\em resistance} associated with the porous network.
%
The figure shows the resistance to decrease rapidly as the bodies erode. Once again the vertical axis is logarithmic, and so the linear trend corresponds to exponential decay of $F_D / (4 \pi \umax)$. We propose that the channelization is responsible for creating such a rapid decline in the network resistance.

@@

% corresponds, pertains, is connected to, involves, is related to, concerns

Comparing the 50 and 6-body simualtions...
Though the area fractions of the two simulations are nearly the same, the 50-body configuration provides about 5 times more resistance.
